---
title: "Week 7 Practice"
author: "Put your name here"
date: "31/10/2019"
output: html_document
---

# Practice 

First, make sure that all the necessary packages are loaded:

```{r}
library(tidyverse)
library(foreign)
```

Here is the breakdown of the tasks we want you to do:

- Read all three datasets in (data_students_1, data_students_2, data_students_3). Since they come in different formats make sure to check your notes from the tutorial.

```{r}
# Read in students_data_1

```

```{r}
# Read in students_data_2

```

```{r}
# Read in students_data_3

```

- Merge the datasets together. First merge data_students_1 and data_students_2, then merge the resulting data with data_students_3. Hint: you will need to use `full_join()`.

```{r}
# Merge students_data_1 and students_data_2 by 'ID'

```

```{r}
# Now merge with students_data_3 (check the solutions for hints - you can use full_join() but you won't need to specify the 'by =' argument)

```

- Work with the final dataset that has information on all 30 students.

```{r}
# Head the data

```

- Provide means, medians and standard deviations for grades in each programme.

```{r}
# Descriptives for all grades

```

- Provide a simple visualisation for each programme.

```{r}
# Grades plot for psych

```

```{r}
# Grades plot for lang

```

```{r}
# Grades plot for phil

```

Now, try to answer the following questions:

- How many students in the dataset received grades above 70?

```{r}
# Filter for grades above 70

```

- What is the mean and median grade for those who got more than 65?

```{r}
# Mean and median for grades above 65

```

- How many students received grades that were between 40 and 50 in philosophy programme?

```{r}
# Phil between 40 and 50

```

- Considering only the philosophy programme, what were the top three grades in the cohort? 

```{r}
# Only phil arranged

```

- Now, for language, what were the three lowest grades in the cohort? 

```{r}
# Only lang arranged

```