---
title: "Week 16"
output: html_document
---

# Walkthrough questions  

## Step.1 Write out the null and alternative hypotheses, and specify our $\alpha$ level.  

Null hypothesis ($H_0$):   
  
Alternative hypothesis ($H_1$):   


We will use $\alpha = ?$  

## Step.2 Get the data!  

We have a sample of 200 rock songs from spotify, available at [https://edin.ac/2wC5rz3](https://edin.ac/2wC5rz3)

```{r message=FALSE}
spotify_songs <- read_csv("https://edin.ac/2wC5rz3")

summary(spotify_songs)
```

## Step.3 Calculate sample mean

```{r}



```


## Step.4 Calculate sample sd

In order to calculate the standard error of the mean, we estimate the population standard deviation ($\sigma$) with the sample standard deviation ($s$).  

```{r}




```

## Step.5 Calculating t

Fill in the blanks:  
$\bar{x}$ = ?
$\mu_0$ = ?
$s$ = ?
$n$ = ?
$SE(\bar{x})$ = ? 

t = ? + ? / ?  

Finally, calculate your t-statistic in R.  

```{r}



```

## New R stuff!  

Going back to the start, we can use our skills with summarise() to calculate all the terms we need for our $t$-statistic: 
```{r}
terms <- spotify_songs %>%
  summarise(
    xbar = mean(tempo),
    s = sd(tempo),
    mu_0 = 125,
    n = n()
  )

terms
```

And then we can plug in these numbers to our equation.    

We now introduce a new technique: the `$` operator. This is similar to the `pull()` function we previously used, as it pulls out the column from a tibble.

```{r}
(terms$xbar - terms$mu_0) / (terms$s / sqrt(terms$n))
```


## Step.6 Critical t values

Using the `qt()` function, calculate the critical value for $\alpha$ = 0.05. This is the smallest *absolute* value of $t$ at which you will reject the null hypothesis.  

+ You'll need to work out the degrees of freedom  
+ You'll also need to think about whether we are performing a two-tailed test or a one-tailed test. If a two-tailed test, then remember that the $\alpha$ is split between the two tails (and so we would reject anything in the most extreme 2.5%)    

```{r}


```


## Step.7 Probability of observed t-statistic  

We have our observed $t$-statistic of 2.1. We know that this is greater than the critical value of 1.65 which means that we have reason to reject the null hypothesis.    

What is the probability of obtaining a $t$-statistic at least as extreme as 2.1, assuming the null hypothesis to be true? In other words, what is the p-value?  

Things you'll need to work out:  

+ Do we want the area under the curve which is to the left or to the right of t = 2.1? 

```{r}



```


# One line of code  

Using our new friend, the `$` operator:
```{r}
t.test(x = spotify_songs$tempo, mu = 125, alternative = "greater")
```

Or, if you prefer:
```{r}
spotify_songs %>% 
   pull(tempo) %>%
   t.test(mu = 125, alternative = "greater")
```


---

# Lab

## Pets' weights  

Data for a sample of 2000 licensed pets from the city of Seattle, USA, can be found at the following url: [https://edin.ac/2VfPzg7](https://edin.ac/2VfPzg7).  
It contains information on the license numbers, issue date and zip-code, as well as data on the species, breeds and weights (in kg) of each pet.  
We are interested in whether the average weight of a dog is different from 20kg.  
  
+ Null hypothesis, $H_0: \mu_1 = 20$  
+ Alternative hypothesis, $H_1: \mu_1 \neq 20$  
  

### Question 1  

Read in the data from the url using `read_csv()`.  
Make sure to *assign* it as an object, using `[yourdataname] <- read_csv(...)` (choose your own name for the data).  
 
Use `summary()` to have a look at your data.  
Which variables are you going to need for our analysis?  
Does anything jump out as relevant?  

```{r}




```


### Question 2

Some of the 2000 pets are not the ones of interest (they aren't dogs).  
Create a new dataset and call it `dogs`, which only has the dogs in it.   

```{r}



```


### Question 3

What does the following command appear to do, and how?  
```{r}
dogs <- dogs %>% filter(!is.na(weight_kg))
```

### Question 4  
Using `summarise()`, calculate $\bar{x}$, $s$ and $n$.  
  
What is $\mu_{0}$, and what are our degrees of freedom ($df$)?  

```{r}



```


### Question 5

Calculate the standardised statistic $t$, using `$` to access the numbers you just calculated above.

```{r}



```

### Question 6

Calculate the p-value using `pt()`.  

```{r}



```


### Question 7  

Finally, use the `t.test()` function.  
Check that the results match the ones you just calculated.  

```{r}



```


## Procrastination scores

The Procrastination Assessment Scale for Students (PASS) was designed to assess how individuals approach decision situations, specifically the tendency of individuals to postpone decisions (see [Solomon & Rothblum, 1984](http://dx.doi.org/10.1037/0022-0167.31.4.503)).  

The PASS assesses the prevalence of procrastination in six areas: writing a paper; studying for an exam; keeping up with reading; administrative tasks; attending meetings; and performing general tasks. 
For a measure of total endorsement of procrastination, responses to 18 questions (each measured on a 1-5 scale) are summed together, providing a single score for each participant (range 0 to 90). 
The mean score from Solomon & Rothblum, 1984 was 33. 

Do Edinburgh University students report endorsing procrastination to less than the norm of 33?  

### Question 8  
Read in the data (a .csv is at [https://edin.ac/2wJgYwL](https://edin.ac/2wJgYwL)), produce some descriptive statistics, and conduct a one sample mean test to evaluate whether Edinburgh University students' average score on the PASS is not equal to 33.  

+ Remember about assumptions of your test!  

```{r}



```

### Question 9  

Write up the results from Question 8.







## Cat weights! 

### Question 10
Without looking at the data (and without googling either), do you think that the average weight of a pet cat is more than/less than 4.5kg?   

Write out your null and alternative hypotheses, and conduct the appropriate test.    

Null hypothesis: ?

Alternative hypothesis: ?

```{r}




```




