---
title: "Week 15: Normal distribution and probability"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# A: Walkthrough

```{r}
library(tidyverse)

# Load the sample
cognitive <- read_csv('https://edin.ac/39jPFa5', col_names = TRUE)

# Display the first six rows
head(cognitive)

# Check the dimensions of the tibble
dim(cognitive)
```

```{r}
ggplot(cognitive, aes(x = CognitiveScore, y = stat(density))) +
  geom_histogram(color = 'white')
```

```{r}
sample_mean <- cognitive %>%
  pull(CognitiveScore) %>%
  mean()

sample_mean
```

```{r}
sample_sd <- cognitive %>%
  pull(CognitiveScore) %>%
  sd()

sample_sd
```

```{r}
ggplot(cognitive) +
  geom_histogram(aes(x = CognitiveScore, y = stat(density)), color = 'white') +
  stat_function(fun = dnorm, 
                args = list(mean = sample_mean, sd = sample_sd), 
                color = 'red', size = 2)
```

```{r}
theory_prob <- pnorm(130, mean = sample_mean, sd = sample_sd) - 
  pnorm(120, mean = sample_mean, sd = sample_sd)
theory_prob
```

```{r}
sample_prob <- cognitive %>%
  summarise(prop = sum(CognitiveScore >= 120 & CognitiveScore <= 130) / n())
sample_prob
```



# B: Lab

```{r, message=FALSE, warning=FALSE}
# Step 1: load the required libraries
library(tidyverse)
library(moderndive)

# Step 2: load the data and select the columns of interest
nfl <- read_tsv('https://edin.ac/2TexAFA')

# Step 3: look at the first rows of the data
head(nfl)

# Step 4: check the dimensions of the tibble
dim(nfl)

# Step 5: select relevant variables
nfl <- nfl %>%
  select(Player, YearlySalary)
```


## Question 1

Compute the sampling distribution of the mean using 10,000 samples of 100 players each.

Display the sampling distribution using a density histogram.
Hint. Remember to use y = stat(density) inside the function aes().


Answer:

```{r}

```


## Question 2
Does the sampling distribution of the mean for samples of size 100 follow a normal curve?


Answer:

...


## Question 3

Find the normal probability density function that models the sampling distribution of the mean for samples of size n = 100.

Plot the normal curve in red on top of the histogram.


Answer:

```{r}

```


## Question 4

What are the population mean and standard deviation?


Answer:

```{r}

```


## Question 5

What are the mean and the standard deviation of the sampling distribution of the mean?

Reminder: Remember that the standard deviation of the sampling distribution of the mean is also known as the standard error (SE) of the mean.


Answer:

```{r}

```


## Question 6

What is the relation between the population mean and the mean of the sampling distribution?


Answer:

...


## Question 7

Do you notice any relation between the population standard deviation and the standard error of the mean?

Hint. If the population standard deviation is $\sigma$, the sampling distribution of the mean should have standard deviation $\frac{\sigma}{\sqrt{n}}$, where $n$ is the sample size.


Answer:

```{r}

```

...


## Question 8: Comparison of the simulation-based and theory-based sampling distribution of the mean

Write down on paper the table below and fill the ? with the mean and the standard error computed using the simulation-based approach and the theory-based approach.

The sampling distribution of the mean is a normal distribution with mean and standard deviation:


Answer:

...


## Question 9

The lengths of human pregnancies are normally distributed with mean $\mu = 266$ days and standard deviation $\sigma = 16$ days.

Plot the normal curve representing the length of human pregnancies.

Hint: This requires the following steps:

1. Find the population mean and standard deviation
2. Create a grid of values for the length of pregnancy variable. Recall from Week 12 that 99.7% of all values lie between the mean plus or minus 3 standard deviations. To be sure to capture all of the values, use the mean plus or minus 4 standard deviations as limits.
3. Compute the normal curve for the grid of values in (2)
4. Plot using geom_line()


Answer:

```{r}

```



## Question 10

What is the proportion of the population with length of pregnancy greater than 280 days?


Answer:

```{r}

```

...


## Question 11
What is the proportion of the population with length of pregnancy between 230 and 260 days?


Answer:

```{r}

```

...



## Question 12

Using the theory-based approach, write down:

1. the mean of the sampling distribution of the average length of pregnancy using samples of size 50;
2. the standard error of the mean.

Plot the sampling distribution of the mean for samples of size 50.

Hint: Remember to create a grid of values going from mean - 4 * SE to mean + 4 * SE.


Answer:

```{r}

```

