---
title: 'Introduction to RMarkdown '
author: "Put your name here"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

Here is the simplified version of what you have seen in the guided part of the lab. Check if you need to update anything above. We will be helping you along with your practice so if anything is unclear just raise your hand :)

To see how changes in your Rmd file affect the final output you can press knit to HTML once you have completed the steps below. 

# Headings

First, lets look at how you would make headings. 

# Section 1 <- `# Section 1`

## Subsection 1 <- `## Subsection 1`

### Sub subsection 1 <- `### Sub subsection 1` 


# Text

You can also vary the format of your text:

`*italics*` returns *italics*
  
`**bold**` returns **bold**
  

# Adding Code

Now, let's add in some code. The syntax for code chunks looks like this:

```{r}
#PUT YOUR CODE HERE
```

Here is my first code chunk:

```{r}
# Create an object a and b
a <- 1
b <- 5

```

We can now use the values assigned in the above chunk like you would in a script:

```{r}
c<-a+b
```

Take a note on what appears in your environment once you have done the above.

Let us make something a bit more meaningful:

```{r}
#Create a variable that stores information on age of participants
age <- c(34, 22, 26, 25, 43,19, 19, 20, 33, 27, 27, 26, 54)

```

Quick plot? We will learn more of these during Week 4.

```{r}
hist(age) # basic quick histogram
```

# Code Chunks edit options

You can edit the way the output gets presented in your final document. 

* `name` - this allows you to name your code chunks but is not necessary.
* `echo` - whether to display the code chunk or just show the results. `echo=FALSE` will embed the code in the document but the reader won't be able to see it.
* `eval` - whether to run the code in the code chunk. `eval=FALSE` will display the code but not run it.
* `warning` - whether to display warning messages in the document.
* `message` - whether to display code messages in the document.
* `results` - whether and how to display the computation of the results.

### Example

### `Warning`
```{r, warning = TRUE, message = TRUE}
library(ggplot2)
```
```{r, warning = FALSE, message = FALSE}
library(ggplot2)
```

You will receive warning here that may say that package was not found. To fix this please run the following in your console:

```{r eval=FALSE}
install.packages('ggplot2)
```

### `Echo`
Now lets try some simple code chunks settings:
```{r, echo = TRUE}
Fab <- 1*5
Fab
```
Same code with `echo = FALSE`:
```{r, echo = FALSE}
Fab <- 1*5
Fab
```

### `Eval`

Here is the same code chunk with `eval = FALSE`. This just shows the code chunk but the code is never run - i.e. you cannot use `Fab` later.

```{r, eval = FALSE}
Fab <- 1*5
Fab
```

### `Results`

Same code chunk but with `results = 'hide'`. Similar to the above, but you CAN use `Fab` later.

```{r, results = 'hide'}
Fab <- 1*5
Fab
```


# Tables

We can create tables that summarise useful information.

Operation       | R code                | Example input    | Example output
-------------   | -------------         |--------------    | -------------
Square root     | `sqrt( )`             | sqrt(100)        | 10
Absolute value  | `abs( )`              | abs(-100)        | 100
Round           | `round(x, digits = )` | round(12.345, 2) | 12.35
Min             | `min(...)`            | min(2.21, 2.22)  | 2.21
Max             | `max(...)`            | max(2.21, 2.22)  | 2.22



          

# Adding links

Remember that you can add links to your text quite easily, using square brackets and including the webpage link e.g. [here] (LINK). In practice, just remove the space.

Can you add a link to the lab notes using the structure above? 

My link is here: [dapR1](PUT LINK HERE)

# Adding figures & pictures

I also want to add a picture to my final document. There is quite nice one that is accessible via this link: http://poster.keepcalmandposters.com/209215.png

Can you include it as well? 

Include a picture from online:
```{r, echo=TRUE,  fig.align="center"}
knitr::include_graphics("PUT LINK TO PICTURE HERE")
```

If you can hide the code from the final output and only display the picture that would be even better! Go back to `Code Chunks` section to see which option you may want to add. Hint: you need to adjust `echo`.


# Soeme more practice

Remember the revenue example we had last week? We practiced using the scripts and the console so it is not that great for us if we want to have it in our notes. Can you incorporate the code and comments into your Rmd file?

 - First create a subheading and call it 'Revenue Example'
 - Fill the code chunk below the relevant information
 

```{r}
#Revenue example
sales <-
royalties <-
revenue <- 
```

Nice! 


You can stop here if you want or if you want to see some more advanced bits, try to work out the example below.

# Inline Text (Advanced)

Inline code allows you to 'fill in' information in your document that is updated as your data is updated. The syntax for inline code starts with `'r` and ends with `'`.

Let's create an example:
```{r}
nextsales <- 20
nextmonth <- "July"
thissales <- 50
thismonth <- "June"
```

Add the inline code::

In **r thismonth**, we sold **r thissales** copies of our Rbook. We are predicted to sell **r nextsales** in **r nextmonth**. This is **r ifelse(nextsales < thissales, "down", "up")** from **r thismonth**.

This will give us the following summary:

In `r thismonth`, we sold `r thissales` copies of our Rbook. We are predicted to sell `r nextsales` in `r nextmonth`. This is `r ifelse(nextsales < thissales, "down", "up")` from `r thismonth`.

Now, compile the file with the information we provided above. Note what you see.

Come back to Rmd editor and adjust the sales to new numbers. Compile again - can you see any changes in your summary paragraph?