# A: Walkthrough

```{r message = FALSE}
library(tidyverse)
library(moderndive)
```

```{r}
outcomes <- tibble(vals = factor(c('S', 'F')))
outcomes
```

```{r}
prob <- c(1/5, 4/5)
samples <- rep_sample_n(outcomes, size = 98, replace = TRUE, reps = 1000, prob = prob)
samples
```

```{r}
null_distribution <- samples %>% 
  group_by(replicate) %>%
  summarise(prop = sum(vals == 'S') / n())

null_distribution
```

```{r}
ggplot(null_distribution, aes(x = prop)) +
  geom_dotplot(binwidth = 0.01, dotsize = 0.5, fill = 'white', stackratio = 0.5) +
  labs(x = expr(hat(p)))
```

```{r}
ggplot(null_distribution, aes(x = prop)) +
  geom_dotplot(binwidth = 0.01, dotsize = 0.5, fill = 'white', stackratio = 0.5) +
  geom_vline(xintercept = 0.26, color = 'tomato1', size = 1) +
  labs(x = expr(hat(p)))
```

```{r}
ggplot(null_distribution, aes(x = prop, fill = (prop >= 0.26))) +
  geom_dotplot(binwidth = 0.01, dotsize = 0.5, stackratio = 0.5) +
  scale_fill_manual(values = c('white', 'tomato1')) +
  labs(x = expr(hat(p)), fill = expr(hat(p) >= 0.26)) +
  theme(legend.position="top")
```

```{r}
pvalue <- null_distribution %>%
  summarise(pvalue = sum(prop >= 0.26) / n())

pvalue
```

```{r}
null_distribution %>%
  summarise(sd = sd(prop))
```



## Summary

Today we have learned to assess how much evidence the sample data bring against the null hypothesis and in favour of the alternative hypothesis.

The null hypothesis, denoted $H_0$, is a claim about a population parameter that is initially assumed to be true. It typically represents "no effect" or "no difference between groups".

The alternative hypothesis, denoted $H_1$, is the claim we seek evidence for.

We assessed the strength of evidence against $H_0$ following these steps:

1. Generate many samples assuming the null hypothesis to be true;
2. Obtain the null distribution by computing the statistic on each of the generated samples;
3. Compute the p-value as the proportion of statistics in the null distribution as extreme or more extreme than the observed statistic, in the direction specified by the alternative hypothesis.

The sample size and the observed statistic in the ESP experiment are:
```{r}
size <- 98
observed_prop <- 0.26
```

We obtained the null distribution with the following code:
```{r}
outcomes <- tibble(vals = factor(c('S', 'F')))
prob <- c(1/5, 4/5)
reps <- 1000

null_distribution <- outcomes %>%
  rep_sample_n(size = size, replace = TRUE, reps = reps, prob = prob) %>%
  group_by(replicate) %>%
  summarise(prop = sum(vals == 'S') / n())
```

The following code chunk plots the null distribution and shows in red the statistics that are as extreme or more extreme than the observed statistic:
```{r}
ggplot(null_distribution, aes(x = prop, fill = (prop >= observed_prop))) +
  geom_dotplot(binwidth = 0.01, dotsize = 0.5, stackratio = 0.5) +
  scale_fill_manual(values = c('white', 'tomato1')) +
  labs(x = expr(hat(p)), fill = expr(hat(p) >= !!observed_prop)) +
  theme(legend.position="top")
```

We compute the p-value as the proportion of statistics in the null distribution which are as extreme or more extreme than the observed statistic:
```{r}
pvalue <- null_distribution %>%
  summarise(pvalue = sum(prop >= observed_prop) / n())

pvalue
```



# B: Lab

Before attempting to answer the following questions, make sure to run the following code chunk, which assumes that you have already installed the packages `tidyverse` and `moderndive`.
If you have not installed them yet, type `install.packages("tidyverse")` and `install.packages("moderndive")` in the R console.

Load the `tidyverse` and `moderndive` packages:
```{r message = FALSE}
library(tidyverse)
library(moderndive)
```

Let's start by loading the data using the function `read_tsv` to read tab separated values:
```{r message=FALSE}
bodytemp <- read_tsv('https://edin.ac/2vkjSaw')
bodytemp
```


## Question 1

1. What are the names of the variables in the dataset?
2. What are the dimensions of the tibble?
3. Are there any missing values? _**Hint:** Use the function `anyNA()` to check if there are any Not Available (NA) entries._

**Solution:**

1. ...
2. ...
3. ...


## Question 2

What is the average temperature in the sample and the standard deviation?

**Solution:**

```{r}

```


## Question 3

State the null and alternative hypothesis for the research question of this lab.

**Solution:**
...


## Question 4

Compute the bootstrap distribution of the mean using 10,000 repeated samples.

Plot the bootstrap distribution using a histogram.

**Solution:**

```{r}

```


## Question 5

Compute the null distribution, which needs to be centred at the value of the parameter specified by the null hypothesis.

Plot the null distribution using a histogram.

_**Hint**: Since you already have the bootstrap distribution, which is centred at the sample mean, you can shift the original sample by adding to each body temperature the difference between the parameter value in the null hypothesis and the sample mean._

**Solution:**

```{r}

```


## Question 6

Show a histogram of the null distribution, superimposing a vertical red line displaying the observed statistic.

**Solution:**

```{r}

```


## Question 7

Calculate the proportion of means as extreme as the observed statistic in the null distribution.

_**Hint:** As this is a double-sided alternative hypothesis, remember to use 2 * p-value of the smallest tail._

**Solution:**

```{r}

```


## Question 8

Using the p-value found in the previous question, how would you answer the research question of the lab?

**Solution:**

...

