---
title: "Week 6 Practice"
author: "Pur Your Name Here"
output: html_document
---

## Practice 
For this week practice we want you to work with simple functions and basic data transformations. 

## Part 1: Work with the  vector of $x$ and apply the following functions.

If we have  $x_1$ represented by:

 $$x_1=\begin{bmatrix}
    -10 \\           
    1\\
    4 \\
    19 \\
    - 6 \\
    0 \\
    14 \\
    -8 \\
    0\\
    12\\
    9\\
    \end{bmatrix} 
    $$

And  $x_2$ represented by:
 $$x_2=\begin{bmatrix}
    -3 \\           
    2\\
    1 \\
    0 \\
    1 \\
    2\\
    3 \\
    5 \\
    -1\\
    -2\\
    -3\\
    \end{bmatrix} 
    $$
    
    
    
Provide the results for:
 
- $y_1=x_1^3$
- $y_2=3x_1-5x_1$
- $y_3=x_1*x_2$


Note: you can use `mutate()` to add an extra variables to represent each of $y_i$. Have a look at the solutions for the hints if not sure where to start.

```{r}
#Create the data with x1 and x2
```


```{r}
#Add three new variables y1, y2 and y3
```


```{r}
#Check what's inside
```


## Part 2: Let's practice with the income data which we worked with last week. 

- Load the data in and work in the template.
- Create an extra variable called **income_thousands** which will provide information on income in thousands of pounds instead of just pounds.
- Visualise the distribution of income. Do you think we need to transform the income variable? Try one of the transformations you have seen today and visualise your data post-transformation (hint: try with log() and mean centering). We have started the steps for you:


```{r}
library(tidyverse)
```


Don't forget to set your working directory.

```{r}
# Read data in
income_london <- read.csv('income_london.csv')
```

```{r}
# Quick description

```


```{r}
#Create variable income_thousands
```


```{r}
#Visualise
```


```{r}
# Create log() transformation and visualise
```


```{r}
# Create mean centered transformation and visualise
```


Take a note of the transformation. Why do you think mean centered income here might be a better transformation of the data compared to that of log()?


## If you finished earlier...
If you finished earlier, go back to your notes online and practice plotting functions by hand and using R (check out the Section 6.6)
