---
title: "Week9_Extra_Prob_Practice"
author: "Data Analysis for Psychology in R"
date: "17/11/2019"
output: html_document
---

## Extra Probability Practice

Work with the distribution we created in the tutorial for guessing on homework quizzes. We want to analyse how likely you are to get specific numbers of question right.

### TRUE/FALSE questions

Work with the TRUE/FALSE example we have seen in the tutorial.
What if the multiple choice had only two choices for the answer (i.e. TRUE or FALSE questions)? The right answer will now have a probability of 0.5 if you were to guess at random. Create a tibble() to show this:

```{r}
homework_guess_true_false <- tibble(right_guess = rbinom(n = 100, 
       size = 10, 
       prob = 0.5)) 
```

### Count the occurencies for right answer

```{r}

```

### Plot the distribution

```{r}

```

Plot with y being a probability (check for hints in the tutorial notes)

```{r}

```

### Use dbinom() to study the probability

Check what the probability is of getting only one question right when guessing at random (use the example in the first chunk to build the rest)

```{r}
dbinom(1, size = 10, prob = 0.5) 
```

Or for five correct answers:

```{r}

```

Or eight: 

```{r}

```

Or put it all together at once (make sure that the probability is 0.5):

```{r}
all_probs <- 
```

### Less than five or more than five?

We can also study what the chances are of getting less than five questions right versus more than five questions right in a TRUE/FALSE setting (check your notes online, you will need `pbinom()`).

```{r}
# Five or less
```

```{r}
# Five or more
```

